taxize - taxonomic search and retrieval in R

### 2013-05-27 inst/doc/taxize_withcode.Rnw
a few changes to ms, made most urls citations

\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
\usepackage[english]{babel}
\usepackage{footnote}
\hypersetup
{
  colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

Taxonomic data is getting easier to obtain through web interfaces (e.g., \url{http://eol.org/}). However, there are a number of good reasons to obtain taxonomic information programatically rather than through a web interface. First, if you have more than a few names to lookup on a website, it can take quite a long time to enter each name, get data, and repeat for each species. Second, programatically getting taxonomic names solves the first problem by looping over a list of names. In addition, doing taxonomic searching, etc. is reproducible. With increasing reports of irreproducibility in science \cite{stodden2010,zimmer2012}, it is extremely important to make science workflows repeatable. Science workflows can now easily incorporate text, code, and images in a single executable document \cite{Yihui2013}.

The R language is the dominant language used by biologists (reference), and now has over 5,000 packages on the R package repository (CRAN) and more than 2,500 packages on other repositories to extend R. R is great for manipulating, visualizing and fitting statistical models to data. However, the key missing piece in R is the ability to get data from the internet within R. Getting data from the web will be increasingly common as more and more data gets moved to the cloud \cite{REF}. Increasingly, data is available from the web via API's, or application programming interfaces. These are bits of code that allow computers to talk to one another using code that is not human readable, but is machine readable. Web APIs often define a number of methods that allow users to search for a species name, or retrieve the synonyms for a species name, for example. A further strength of APIs is that they are language agnostic, meaning that data can be consumed in almost any computing context, allowing users to interact with the web API without having to know the details of the code. Whereas, if data are stored in an Excel file, for example, the file can only be opened in a few programs.

@@

There are a large suite of applications developed around the problem of searching for, resolving, and getting higher taxonomy for species names. For example, \textbf{Linnaeus} \url{http://linnaeus.sourceforge.net/} provides the ability to search for taxonomic names in documents and normalize those names found. In addition, there are many web interfaces to search for and normalize names such as Encyclopedia of Life's Global Names Resolver \url{http://resolver.globalnames.org/}, uBio tools \url{http://www.ubio.org/index.php?pagename=sample_tools}, and iPlant's Taxonomic Name Resolution Service \url{http://tnrs.iplantcollaborative.org/}.

@@

One data source available in taxize is available in another R pacakge that solely interact with that data source. We provide a few convenience functions that wrap functions in \emph{taxonstand}, an R interface to The Plant List \url{http://www.theplantlist.org}.

@@

\subsection{Installing taxize}

First, we must install taxize. There are two versions of taxize, a) a stable release that can be installed from the R package repository, CRAN, and b) from GitHub \url{https://github.com/ropensci/taxize_}, where the code is developed.

@@

<<installtaxize, eval=FALSE>>=
## From CRAN
install.packages("taxize")

@@

This is a common task in biology. We often have a list of species names and we want to know if a) we have the most up to date names, b) our names are spelled correctly, and c) if we have common names, we likely need the scientific names. One way to resolve names is via the Global Names Resolver (GNR) service provided by the Encyclopedia of Life (\url{http://resolver.globalnames.org/}).

@@

It looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}. Another approach uses the Taxonomic Name Resolution Service via the Taxosaurus API (\url{http://taxosaurus.org/}) developed by iPLant and the Phylotastic organization.

@@

<<interactive, eval=FALSE, echo=TRUE, cache=TRUE>>=
# In this example, this search term has many matches. The function returns a data.frame of the matches, and asks for user input. Enter the row number of the name you want. 
get_tsn(searchterm="Elisabethae", searchtype="sciname")

@@

There are currently no resources for getting a phylogeny of animals simply from species names. However, a few projects are working on this problem, including the Open Tree of Life \url{URL}. We will incorporate these resources when the appropriate APIs are available.

@@

If you aren't a taxonomic specialist on a particular taxon you likely don't know what children taxa are within a family, or within a genus. This task becomes especially unwieldy when there are a large number of taxa downstream. You can of course go to a website like Wikispecies (\url{http://species.wikimedia.org/wiki/Main_Page}) or Encyclopedia of Life (\url{http://eol.org/}) to get downstream names. However, taxize provides an easy way to programatically search for downstream taxa, both for the Catalogue of Life (CoL; \url{http://www.catalogueoflife.org/}) and the Integrated Taxonomic Information System (\url{http://www.itis.gov/}). Here is a short example using the CoL in which we want to find all the species within the genus \emph{Apis} (honey bees).

@@

taxize is currently under development and will be for some time given the large number of data sources kitted together in the package, and the fact that APIs for each data source can change, requiring changes in taxize code. Contributions to taxize are strongly encouraged, and can be easily done using GitHub here \url{http://github.com/ropensci/taxize_}. We hope taxize will be taking up by the community and developed collaboratively, making it progressively better through time as new use cases arise, bug reports are squashed, and pull requests are merged.

@@

\begin{itemize}
    \item{Appendix A.} A complete reproducible workflow, from a species list to a phylogeny. 
\end{itemize}

>>>>

% \usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
% \usepackage[english]{babel}
\usepackage{footnote}
\hypersetup
{
  colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

Taxonomic data is getting easier to obtain through web interfaces (e.g., \cite{eol}). However, there are a number of good reasons to obtain taxonomic information programatically rather than through a web interface. First, if you have more than a few names to lookup on a website, it can take quite a long time to enter each name, get data, and repeat for each species. Second, programatically getting taxonomic names solves the first problem by looping over a list of names. In addition, doing taxonomic searching, etc. is reproducible. With increasing reports of irreproducibility in science \cite{stodden2010,zimmer2012}, it is extremely important to make science workflows repeatable. Science workflows can now easily incorporate text, code, and images in a single executable document \cite{yihui2013}.

The R language is the dominant language used by biologists (reference), and now has over 5,000 packages on the R package repository (CRAN) and more than 2,500 packages on other repositories to extend R. R is great for manipulating, visualizing and fitting statistical models to data. However, the key missing piece in R is the ability to get data from the internet within R. Getting data from the web will be increasingly common as more and more data gets moved to the cloud. Increasingly, data is available from the web via API's, or application programming interfaces. These are bits of code that allow computers to talk to one another using code that is not human readable, but is machine readable. Web APIs often define a number of methods that allow users to search for a species name, or retrieve the synonyms for a species name, for example. A further strength of APIs is that they are language agnostic, meaning that data can be consumed in almost any computing context, allowing users to interact with the web API without having to know the details of the code. Whereas, if data are stored in an Excel file, for example, the file can only be opened in a few programs.

@@

There are a large suite of applications developed around the problem of searching for, resolving, and getting higher taxonomy for species names. For example, Linnaeus \cite{linnaeus} provides the ability to search for taxonomic names in documents and normalize those names found. In addition, there are many web interfaces to search for and normalize names such as Encyclopedia of Life's Global Names Resolver \cite{eolgnr}, uBio tools \cite{ubio}, and iPlant's Taxonomic Name Resolution Service \cite{tnrs}.

@@

One data source available in taxize is available in another R pacakge that solely interact with that data source. We provide a few convenience functions that wrap functions in \emph{taxonstand}, an R interface to The Plant List \cite{theplantlist}.

@@

\subsection*{First, install taxize}

First, we must install taxize. There are two versions of taxize, a) a stable release that can be installed from the R package repository, CRAN, and b) from GitHub \cite{github_taxize}, where the code is developed.

@@

<<installtaxize, eval=TRUE>>=
## From CRAN
install.packages("taxize")

@@

This is a common task in biology. We often have a list of species names and we want to know if a) we have the most up to date names, b) our names are spelled correctly, and c) if we have common names, we likely need the scientific names. One way to resolve names is via the Global Names Resolver (GNR) service provided by the Encyclopedia of Life \cite{eolgnr}.

@@

It looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}. Another approach uses the Taxonomic Name Resolution Service via the Taxosaurus API (\cite{taxosaurus}) developed by iPLant and the Phylotastic organization.

@@

<<interactive, eval=TRUE, echo=TRUE, cache=TRUE>>=
# In this example, this search term has many matches. The function returns a data.frame of the matches, and asks for user input. Enter the row number of the name you want. 
get_tsn(searchterm="Elisabethae", searchtype="sciname")

@@

There are currently no resources for getting a phylogeny of animals simply from species names. However, a few projects are working on this problem, including the Open Tree of Life \cite{opentree}. We will incorporate these resources when the appropriate APIs are available.

@@

If you aren't a taxonomic specialist on a particular taxon you likely don't know what children taxa are within a family, or within a genus. This task becomes especially unwieldy when there are a large number of taxa downstream. You can of course go to a website like Wikispecies (\cite{wikispecies}) or Encyclopedia of Life (\cite{eol}) to get downstream names. However, taxize provides an easy way to programatically search for downstream taxa, both for the Catalogue of Life (CoL; \cite{col}) and the Integrated Taxonomic Information System (\cite{itis}). Here is a short example using the CoL in which we want to find all the species within the genus \emph{Apis} (honey bees).

@@

taxize is currently under development and will be for some time given the large number of data sources kitted together in the package, and the fact that APIs for each data source can change, requiring changes in taxize code. Contributions to taxize are strongly encouraged, and can be easily done using GitHub here \cite{github_taxize}. We hope taxize will be taking up by the community and developed collaboratively, making it progressively better through time as new use cases arise, bug reports squashed, and contributions merged.

@@

\begin{itemize}
    \item{Appendix A.} A complete reproducible workflow, from a species list to a phylogeny, and a species distribution map. 
\end{itemize}

### 2013-05-27 inst/doc/taxize_withcode.tex
a few changes to ms, made most urls citations

\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
\usepackage[english]{babel}
\usepackage{footnote}
\hypersetup
{
  colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{## From CRAN}
\hlfunctioncall{install.packages}(\hlstring{"taxize"})

\hlcomment{## From GitHub}
\hlfunctioncall{install.packages}(\hlstring{"devtools"})
\hlfunctioncall{require}(devtools)
\hlfunctioncall{install_github}(\hlstring{"taxize_"}, \hlstring{"ropensci"})
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\hlcomment{# Let's take a peek at the data, excluding the data source ID and score}
\hlcomment{# columns}
temp[, -\hlfunctioncall{c}(1, 4)]
\end{alltt}
\end{kframe}
\end{knitrout}

It looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}. Another approach uses the Taxonomic Name Resolution Service via the Taxosaurus API (\url{http://taxosaurus.org/}) developed by iPLant and the Phylotastic organization.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
specieslist <- \hlfunctioncall{c}(\hlstring{"Abies procera"}, \hlstring{"Pinus contorta"})
\hlfunctioncall{classification}(\hlfunctioncall{get_tsn}(specieslist, \hlstring{"sciname"}))
\end{alltt}
\begin{verbatim}
[[1]]
        parentName parentTsn      rankName       taxonName    tsn
1                                  Kingdom         Plantae 202422
2          Plantae    202422    Subkingdom  Viridaeplantae 846492
3   Viridaeplantae    846492  Infrakingdom    Streptophyta 846494
4     Streptophyta    846494      Division    Tracheophyta 846496
5     Tracheophyta    846496   Subdivision Spermatophytina 846504
6  Spermatophytina    846504 Infradivision    Gymnospermae 846506
7     Gymnospermae    846506         Class       Pinopsida 500009
8        Pinopsida    500009         Order         Pinales 500028
9          Pinales    500028        Family        Pinaceae  18030
10        Pinaceae     18030         Genus           Abies  18031
11           Abies     18031       Species   Abies procera 181835

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# Using ITIS}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"itis"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\begin{alltt}

\hlcomment{# Using NCBI}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"ncbi"})
\end{alltt}
\begin{verbatim}
      family
1 Asteraceae
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# In this example, this search term has many matches. The function returns}
\hlcomment{# a data.frame of the matches, and asks for user input. Enter the row}
\hlcomment{# number of the name you want.}
\hlfunctioncall{get_tsn}(searchterm = \hlstring{"Elisabethae"}, searchtype = \hlstring{"sciname"})

\hlcomment{# You can pass in a long list of taxonomic names}
splist <- \hlfunctioncall{c}(\hlstring{"annona cherimola"}, \hlstring{"annona muricata"}, \hlstring{"quercus robur"}, \hlstring{"shorea robusta"}, 
    \hlstring{"pandanus patina"}, \hlstring{"oryza sativa"}, \hlstring{"durio zibethinus"})
\hlfunctioncall{get_tsn}(searchterm = splist, searchtype = \hlstring{"sciname"})

\hlcomment{# In this example, note that no match at all returns an NA}
\hlfunctioncall{get_uid}(sciname = \hlfunctioncall{c}(\hlstring{"Chironomus riparius"}, \hlstring{"aaa vva"}))
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\hlcomment{# Fetch the tree - the formatting of names and higher taxonmy is done}
\hlcomment{# within the function}
tree <- \hlfunctioncall{phylomatic_tree}(taxa = taxa, get = \hlstring{"POST"}, informat = \hlstring{"newick"}, method = \hlstring{"phylomatic"}, 
    storedtree = \hlstring{"R20120829"}, taxaformat = \hlstring{"slashpath"}, outformat = \hlstring{"newick"}, 
    clean = \hlstring{"true"})

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
ia <- \hlfunctioncall{iucn_summary}(\hlfunctioncall{c}(\hlstring{"Panthera uncia"}, \hlstring{"Lynx lynx"}))
\hlfunctioncall{iucn_status}(ia)
\end{alltt}
\begin{verbatim}
Panthera uncia      Lynx lynx 
          "EN"           "LC" 
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\hlcomment{# Does the string 'RAG1' exist in any of the gene names?}
out[\hlfunctioncall{grep}(\hlstring{"RAG1"}, out$genesavail, ignore.case = T), ]
\end{alltt}
\begin{verbatim}
        spused length
414 Umbra limi    732
427 Umbra limi    959
434 Umbra limi   1631
                                                                            genesavail
414 isolate UlimA recombinase activating protein 1 (rag1) gene, exon 3 and partial cds
427           recombination-activating protein 1 (RAG1) gene, intron 2 and partial cds
434                        recombination-activating protein 1 (RAG1) gene, partial cds
    access_num       ids
414   JX190826 394772608
427   AY459526  45479841
434   AY380548  38858304
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

% \usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
% \usepackage[english]{babel}
\usepackage{footnote}
\hypersetup
{
  colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{## From CRAN}
\hlfunctioncall{install.packages}(\hlstring{"taxize"})
\end{alltt}
\begin{verbatim}

The downloaded binary packages are in
	/var/folders/kv/rpbb_gw965l0lhrq8w92n8j40000gn/T//Rtmp44FcYx/downloaded_packages
\end{verbatim}
\begin{alltt}

\hlcomment{## From GitHub}
\hlfunctioncall{install.packages}(\hlstring{"devtools"})
\end{alltt}
\begin{verbatim}

The downloaded binary packages are in
	/var/folders/kv/rpbb_gw965l0lhrq8w92n8j40000gn/T//Rtmp44FcYx/downloaded_packages
\end{verbatim}
\begin{alltt}
\hlfunctioncall{require}(devtools)
\hlfunctioncall{install_github}(\hlstring{"taxize_"}, \hlstring{"ropensci"})
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\hlcomment{# Let's take a peek at the data, excluding the data source ID and score}
\hlcomment{# columns}
temp[, -\hlfunctioncall{c}(1, 4)]
\end{alltt}
\begin{verbatim}
    submitted_name                 name_string                   title
1 Helianthos annus        Helianthus annuus L.       Catalogue of Life
3 Helianthos annus            Helianthus annus GBIF Taxonomic Backbone
4 Helianthos annus            Helianthus annus                     EOL
5 Helianthos annus         Helianthus annus L.                     EOL
6 Helianthos annus            Helianthus annus           uBio NameBank
2    Homo saapiens Homo sapiens Linnaeus, 1758       Catalogue of Life
\end{verbatim}
\end{kframe}
\end{knitrout}

It looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}. Another approach uses the Taxonomic Name Resolution Service via the Taxosaurus API (\cite{taxosaurus}) developed by iPLant and the Phylotastic organization.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
specieslist <- \hlfunctioncall{c}(\hlstring{"Abies procera"}, \hlstring{"Pinus contorta"})
\hlfunctioncall{classification}(\hlfunctioncall{get_tsn}(specieslist, \hlstring{"sciname"}))
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Abies procera '

Retrieving data for species ' Pinus contorta '
[[1]]
        parentName parentTsn      rankName       taxonName    tsn
1                                  Kingdom         Plantae 202422
2          Plantae    202422    Subkingdom  Viridaeplantae 846492
3   Viridaeplantae    846492  Infrakingdom    Streptophyta 846494
4     Streptophyta    846494      Division    Tracheophyta 846496
5     Tracheophyta    846496   Subdivision Spermatophytina 846504
6  Spermatophytina    846504 Infradivision    Gymnospermae 846506
7     Gymnospermae    846506         Class       Pinopsida 500009
8        Pinopsida    500009         Order         Pinales 500028
9          Pinales    500028        Family        Pinaceae  18030
10        Pinaceae     18030         Genus           Abies  18031
11           Abies     18031       Species   Abies procera 181835

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# Using ITIS}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"itis"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Helianthus annuus '
[1] "Asteraceae"
\end{verbatim}
\begin{alltt}

\hlcomment{# Using NCBI}
\hlfunctioncall{tax_name}(query = \hlstring{"Helianthus annuus"}, get = \hlstring{"family"}, db = \hlstring{"ncbi"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Helianthus annuus '
[1] "Asteraceae"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# In this example, this search term has many matches. The function returns}
\hlcomment{# a data.frame of the matches, and asks for user input. Enter the row}
\hlcomment{# number of the name you want.}
\hlfunctioncall{get_tsn}(searchterm = \hlstring{"Elisabethae"}, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Elisabethae '

                           combinedname    tsn
1                  Bittacus elisabethae 810495
2            Caenaugochlora elisabethae 754498
3                  Ceratina elisabethae 765455
4            Dolicheremaeus elisabethae 740429
5                  Euphione elisabethae  64893
6                 Harmothoe elisabethae 573734
7                  Hydraena elisabethae 815669
8                    Lizzia elisabethae  48875
9                   Melinna elisabethae  67765
10            Myrmicocrypta elisabethae 580527
11           Myzomela boiei elisabethae 713955
12             Notosacantha elisabethae 838430
13               Odontomyia elisabethae 627043
14             Parabelbella elisabethae 737749
15                Peltigera elisabethae 191230
16          Polymorphanisus elisabethae 600292
17        Pseudopterogorgia elisabethae  52264
18                  Rhantus elisabethae 815107
19              Scyllarides elisabethae 553004
20       Sellnickochthonius elisabethae 735137
21       Serilophus lunatus elisabethae 711636
22                Stenhelia elisabethae  87927
23              Tetramorium elisabethae 582397
24                Theridion elisabethae 868794
25             Trichoniscus elisabethae 597365
26 Trichoniscus elisabethae estoniensis 598124

More than one TSN found for species ' Elisabethae '!

          Enter rownumber of species (other inputs will return 'NA'):

Returned 'NA'!
 1 
NA 
attr(,"class")
[1] "tsn"
\end{verbatim}
\begin{alltt}

\hlcomment{# You can pass in a long list of taxonomic names}
splist <- \hlfunctioncall{c}(\hlstring{"annona cherimola"}, \hlstring{"annona muricata"}, \hlstring{"quercus robur"}, \hlstring{"shorea robusta"}, 
    \hlstring{"pandanus patina"}, \hlstring{"oryza sativa"}, \hlstring{"durio zibethinus"})
\hlfunctioncall{get_tsn}(searchterm = splist, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' annona cherimola '

Retrieving data for species ' annona muricata '

Retrieving data for species ' quercus robur '

Retrieving data for species ' shorea robusta '

Retrieving data for species ' pandanus patina '

Retrieving data for species ' oryza sativa '

Retrieving data for species ' durio zibethinus '
[1] "506198" "18098"  "19405"  "506787" "507376" "41976"  "506099"
attr(,"class")
[1] "tsn"
\end{verbatim}
\begin{alltt}

\hlcomment{# In this example, note that no match at all returns an NA}
\hlfunctioncall{get_uid}(sciname = \hlfunctioncall{c}(\hlstring{"Chironomus riparius"}, \hlstring{"aaa vva"}))
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Chironomus riparius '

Retrieving data for species ' aaa vva '
[1] "315576" NA      
attr(,"class")
[1] "uid"
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\hlcomment{# Fetch the tree - the formatting of names and higher taxonmy is done}
\hlcomment{# within the function}
tree <- \hlfunctioncall{phylomatic_tree}(taxa = taxa, get = \hlstring{"POST"}, informat = \hlstring{"newick"}, method = \hlstring{"phylomatic"}, 
    storedtree = \hlstring{"R20120829"}, taxaformat = \hlstring{"slashpath"}, outformat = \hlstring{"newick"}, 
    clean = \hlstring{"true"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Poa annua '

Retrieving data for species ' Abies procera '

Retrieving data for species ' Helianthus annuus '
\end{verbatim}
\begin{alltt}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
ia <- \hlfunctioncall{iucn_summary}(\hlfunctioncall{c}(\hlstring{"Panthera uncia"}, \hlstring{"Lynx lynx"}))
\hlfunctioncall{iucn_status}(ia)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\hlcomment{# Does the string 'RAG1' exist in any of the gene names?}
out[\hlfunctioncall{grep}(\hlstring{"RAG1"}, out$genesavail, ignore.case = T), ]
\end{alltt}
\begin{verbatim}
        spused length
413 Umbra limi    732
427 Umbra limi    959
434 Umbra limi   1631
                                                                            genesavail
413 isolate UlimA recombinase activating protein 1 (rag1) gene, exon 3 and partial cds
427           recombination-activating protein 1 (RAG1) gene, intron 2 and partial cds
434                        recombination-activating protein 1 (RAG1) gene, partial cds
    access_num       ids
413   JX190826 394772608
427   AY459526  45479841
434   AY380548  38858304
\end{verbatim}
\end{kframe}
\end{knitrout}
